Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-23 supplement.tex

@@

\begin{figure}
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering\includegraphics[width=0.9\textwidth]{Figs/td-widths}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-widths}
\end{figure}

\begin{figure}
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  \includegraphics[width=0.9\textwidth]{Figs/td-times}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-times}
\end{figure}

\begin{figure}[t!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  \verbatiminput{Figs/f3_100_0.txt}
  \includegraphicstop[width=0.6\textwidth]{Figs/td-example-basepair}%
  \includegraphicstop[width=0.35\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{fig:td-example}
\end{figure}

@@

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 20.000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{fig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the runtime, as noticeable in Figure~\ref{fig:run-times-redprint}.

@@

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 20.000 sequences. }
  \label{fig:run-times-redprint}
\end{figure}
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 20.000 sequences.}
  \label{fig:run-times-rbp}
\end{figure}

@@

\begin{figure}[tb]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, resepectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist}
\end{figure}

\begin{figure}
  \begin{center}
    {\sf \bfseries A}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/PKB00211_PKB00239_0_GC_content}
    ~~~~~{\sf \bfseries B}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/supp_3str_GC_content}
  \end{center}
  \caption{Distribution of the \GCb-content \GCb\% in the experiments of (A) Fig.~\ref{fig:energydist-pk} of the main text and (B) Suppl.~Fig.~\ref{fig:energydist}.}
  \label{fig:gc-content}
\end{figure}

@@

\begin{figure}
  \centering
  A)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{fig:training-cor}
\end{figure}

>>>>

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering\includegraphics[width=0.9\textwidth]{Figs/td-widths}
  \caption{Treewidths for multi-target RNA design instances of
    different complexity. Distributions of treewidths shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-widths}
\end{figure}

\begin{figure}[h!]
  %\hspace*{2cm}\textbf{Base pair model}\hspace{4.75cm}\textbf{Stacking model}\\[-28pt]
  \centering
  \includegraphics[width=0.9\textwidth]{Figs/td-times}
  \caption{Computation time of tree decompositions for
    multi-target RNA design instances of different complexity.
    Distributions of times (in ms/instance) shown as boxplots for the base pair (left) and stacking model (right).}
  \label{fig:td-times}
\end{figure}

@@

\begin{figure}[h!]
  %% the figure shows benchmark_inputs/RNAfold/3str/f3.100.0.inp
  \centering
  \verbatiminput{Figs/f3_100_0.txt}
  \includegraphicstop[width=0.6\textwidth]{Figs/td-example-basepair}%
  \includegraphicstop[width=0.35\textwidth]{Figs/td-example-stacking}
  \caption{Example instance from RF3 (top) with its tree decompositions in the base pair (left) and stacking model (right). The respective treewidths are 2 and 12.}
  \label{fig:td-example}
\end{figure}

@@

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 10000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{fig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the runtime, as noticeable in Figure~\ref{fig:run-times-redprint}.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences. }
  \label{fig:run-times-redprint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.45\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences.}
  \label{fig:run-times-rbp}
\end{figure}

@@

\begin{figure}[h!]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, resepectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    {\sf \bfseries A}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/PKB00211_PKB00239_0_GC_content}
    ~~~~~{\sf \bfseries B}~\includegraphicstop[width=0.45\textwidth]{Figs/Plots/supp_3str_GC_content}
  \end{center}
  \caption{Distribution of the \GCb-content \GCb\% in the experiments of (A) Fig.~\ref{fig:energydist-pk} of the main text and (B) Suppl.~Fig.~\ref{fig:energydist}.}
  \label{fig:gc-content}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  A)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/basepaircor}
  B)\includegraphicstop[width=0.4\textwidth,trim=0 0 0 50,clip]{Figs/stackingcor}
  \caption{Validation of the trained parameters for the A) base pair
    model B) stacking model for predicting energies of the
    independently sampled sequences in the test set. We show
    correlation plots against the Turner energies reported by the
    ViennaRNA package.}
  \label{fig:training-cor}
\end{figure}
